---
title: "Supper_store_IP"
author: "Stephen Njuguna"
date: "2/4/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Supper Market, Marketing Strategy 

## Defining the question 

Coming up with the best marketing strategy 

## Defining the metric for success

Being able to inform the marketing the department of the best marketing the strategy 
that will result to the highest number of sales.

## Understanding the context

Marketing strategy provides an organization an edge over it's competitors. Strategy helps in developing goods and services with best profit making potential. It also helps an organization to make optimum utilization of its resources so as to provide a sales message to it's target market and increase its profit margins.

## Experimental design

  * Data Cleaning and Preparation
  * Deal with missing values
  * Deal with duplicated records
  * Univariate Analysis
  * Bivariate Analysis
  * Multivariate Analysis
  * Dimension Reduction
  * Feature Selection
  * Association Analysis
  * Anomaly Reduction 
  * Conclusion
  * Recommendations 


## Reading data 

```{r}
#Loading dataset 
supper_df <- read.csv('http://bit.ly/CarreFourDataset')
head(supper_df)
```
```{r}
#Preview the data types  

str(supper_df)

```

```{r}
#Preview dataset dimension 
dim(supper_df)
```
## Data Cleaning 

```{r}
#Checking for missing values
colSums(is.na(supper_df))
```

```{r}
#Checking for duplicates and dealing with them 
supper_df <- unique(supper_df)

```



## Exploratory Data Analysis

### Univeriate Analysis 

```{r}
#Previewing statistical analysis of the dataset 
numeric_supper_df <- supper_df[c('Unit.price','Quantity','Tax','cogs','gross.margin.percentage','gross.income','Rating','Total')]

summary(numeric_supper_df)

```

```{r}
#Plot customer type
table(supper_df$Payment)
barplot(table(supper_df$Payment))
```

Most customers are paying with cash and Ewallet


```{r}
#Plot products that most customers by 
table(supper_df$Product.line)
barplot(table(supper_df$Product.line))
```


Fashion accessories are the most bought products followed closely by food and beverages,
the least bought products are health and beauty.

## Biavariate Analysis


```{r}
#Plot stacked bar plot of branch and product sales 
library(ggplot2)
ggplot(supper_df, aes(fill=supper_df$Product.line, y=supper_df$Total, x=supper_df$Branch)) + 
    geom_bar(position="fill", stat="identity")


```

Some observations :
  * All the branch have equal sale of electronic accessories.
  * Branch c has highest sale of fashion accessories.
  * Branch c has highest sales of food and beverages.
  * Branch A has highest sale of home and lifestyle compared to the others.


```{r}
#Plot bar plot of branch and and rate

rate <- aggregate(supper_df$Rating, list(supper_df$Branch), FUN=mean)

barplot(x ~ Group.1,data = rate)
```


Branch B of the supper market has the least rating 


```{r}
#Plot bar plot of branch and and total sales

total <- aggregate(supper_df$Total, list(supper_df$Branch), FUN=mean)

barplot(x ~ Group.1,data = total)
```

Branch A of the supper market has the least sales

```{r}
#Plot scatter plot of total sales and rating 

plot(supper_df$Total, supper_df$Rating, xlab="Total sales", ylab="Rating")
```

There is a high correlation between between sales and rating.


## Multivariate Analysis

### Dimension Reduction 

**t-Distributed Stochastic Neighbor Embedding (t-SNE)**

```{r}
#Label Encoding Categorical to numeric data
supper_df$Branch <- as.numeric(factor(supper_df$Branch))
supper_df$Customer.type <- as.numeric(factor(supper_df$Customer.type))
supper_df$Gender <- as.null(factor(supper_df$Gender))
supper_df$Product.line <- as.numeric(factor(supper_df$Product.line))
supper_df$Payment <- as.numeric(factor(supper_df$Payment))
```


```{r}
#Selecting the numeric data
library(dplyr)
new_df <- select_if(supper_df, is.numeric)

```


```{r}
#Create t-sne model 
library(Rtsne)
# Executing the algorithm 
tsne <- Rtsne(new_df, dims = 2, perplexity=30, verbose=TRUE, max_iter = 500)

# Getting the duration of execution
exeTimeTsne <- system.time(Rtsne(new_df, dims = 2, perplexity=30, verbose=TRUE, max_iter = 500))
```

```{r}
# Plotting our graph and closely examining the graph
plot(tsne$Y, t='n', main="tsne")
text(tsne$Y)

```

### Feature selection 

**Filter Methods**

```{r}
#Import Libraries
library(corrplot)
library(caret)

# Calculating the correlation matrix
cor_Matrix <- cor(new_df)
# Find attributes that are highly correlated

high_Correlated <- findCorrelation(cor_Matrix, cutoff=0.75)
```

```{r}
# Show the highly correlated attributes

high_Correlated

names(new_df[,high_Correlated])
```


Tax, Cogs and Gross income are the attributes that are highly correlated.


## Assocaition Analysis

```{r}
# Loading the arules library
library(arules)
#Loading the dataset 
trans_df <- read.transactions('http://bit.ly/SupermarketDatasetII',sep = ',',rm.duplicates = TRUE)

#Inspect transaction class

class(trans_df)
```

```{r}
#Inspect the top transactions
inspect(trans_df[1:5])
```

```{r}
#Previewing the summary of the transaction 
summary(trans_df)
```

Mineral water and eggs are among the top purchased items from the supermarket 


```{r}
# Displaying top 10 most common items in the transactions data 
# and the items whose relative importance is at least 10%

par(mfrow = c(1, 2))

# plot the frequency of items
itemFrequencyPlot(trans_df, topN = 10,col="darkgreen")
itemFrequencyPlot(trans_df, support = 0.1,col="darkred")
```

Mineral water ,eggs and spaghetti are among the most bought items in the supermarket
but also are among the items that have less that 10% importance.


```{r}
#Creating the association analysis(Apriori model) model 
rule <- apriori (trans_df, parameter = list(supp = 0.001, conf = 0.8))
rule
```

With a confidence of 0.8 and association of 0.001 the model as 74 rules.


```{r}
#Previewing the summary of the rules 
summary(rule)
```

```{r}
#Observing some of the rules built by the model 
inspect(rule[1:5])
```


## Anomaly Detection 

### Tidy Anomaly Detection

```{r}
#Import library 
library(anomalize) 
library(tidyverse)
```


```{r}
#Load dataset 

library(tibbletime)
sales<- read.csv('http://bit.ly/CarreFourSalesDataset')
sales <- unique(sales)
#head(sales)
#Converting column Date to Date type 
sales$Date <- as.Date(sales$Date , format = '%m/%d/%Y')

#Data preprocessing converting data frame to tibble 
sales_tb <-  sales %>%as_tbl_time(Date)

#Anomaly detection model
sales_tb %>%
  group_by(Date) %>%
  summarise(Orders = sum(Sales)) %>%
  time_decompose(Orders, method = "twitter", frequency = "auto", trend = "auto") %>%
  anomalize(remainder,method = "gesd", alpha = 0.05, max_anoms = 0.2) %>%
  time_recompose() %>%
  # Anomaly Visualization
  plot_anomaly_decomposition()

```


## Conclusion 

  * Fashion accessories are the most bought products followed closely by food and beverages,
  the least bought are health and beauty products.
  * All the branch have equal sale of electronic accessories.
  * Branch c has highest sale of fashion accessories.
  * Branch c has highest sales of food and beverages.
  * Branch A has highest sale of home and lifestyle compared to the others.
  * Branch A has the least total sales 
  * Branch B has the least rating 
  * The is a strong positive correlation between sales and rating 
  
  
## Recommendations 

  * Improve customer service in branch B, this will improve total sales as we have
  seen that is a strong positive correlation between sales and ratings.
  * Since there is big percentage of customers paying using E-wallet, you can run 
  online advertisements using the association analysis model.
  * Put more resources on each branch on the products that they least sale.Run more promotions 
  on those products. 
  * Put more marketing resources on branch A, since it has the least sales. 
  









